The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2018-08-28 GRB111117A_old.tex
Published

\title{The host galaxy of the short GRB~111117A at $z = 2.211$: impact on the short GRB redshift distribution and progenitor channels\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}

@@

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

@@

The electromagnetic signals from sGRBs are a promising channel to accurately
localize NS mergers, which holds the promise for a detection of an associated
gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
of a sGRB and a GW will provide new promising ways to constrain the binary
inclination angle \citep{Arun2014} and measure cosmological distances
\citep{Nissanke2010}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned down to 3 \AA~per pixel for presentation purposes. Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum. The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
spectrum in Fig.~\ref{fig:SED}.

@@

The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
profiles shows an asymmetric "double-horned" profile, indicating that we are
seeing a galaxy with a large degree of coherent rotational motion relative to
the line-of-sight. If we assume that we are viewing a spiral galaxy edge-on,
this is a measure of the rotational velocity of the gas. If we assume that the
spectral resolution and the turbulent width of the lines are negligible compared
to the rotational velocity, we can, based on the projected size of the source
and the width of the lines, put a constraint on the dynamical mass of the galaxy
\citep{DeBlok2014}. Based on the physical size along the slit and the velocity
width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
M$_\odot$. Because we are viewing the host inclined at an angle relative to
edge-on and because the slit is not aligned along the long axis of the host,
this value is a lower limit.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
Sect. \ref{xray}, there is likely no afterglow contamination on the measured
photometry.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X-ray temporal and spectral analysis}\label{xray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The redshift evolution of $N_H$ in the hosts of lGRBs is not reproduced by
\citet{Buchner2017}, using a different $N_H$ inference methodology. Instead a
correlation between $N_H$ and host stellar mass is suggested. Assuming that the
different $N_H$-fitting methodologies yield comparable results, GRB~111117A is
an outlier from the relation suggested by \citet{Buchner2017} by more than the
intrinsic scatter, although some lGRB hosts populate a similar region in the
$N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
surface luminosity at the explosion site \citep{Lyman2017}. The large offset of
GRB~111117A relative to the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013} is difficult to reconcile with galaxy-scale gas providing
the X-ray absorption. Along with the absence of dust, the large offset from the
host center indicates that the high $N_H$ arises because the density in the GRB
surrounding is high, or because the light from the afterglow traverses a region
of dense gas. The optical darkness of the GRB is an additional hint of the high
density in the GRB surroundings or along the line of sight.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs. The detections are replaced with contours and the limits are shown with arrows. Marginalizations over both axes are shown where the limits are shown as transparent bars and detections as solid ones.}
	\label{fig:NH_z}
\end{figure}

@@

If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed
60 per cent of the redshifts (6 out of 10 hosts), due too the host being too
faint, were they at the redshift of GRB~11117A. The corresponding number is
around 45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
the average SFR of galaxies hosting lGRBs is higher than for galaxies hosting
sGRBs, the fraction of missed burst redshifts is likely higher although the
cosmic SFR evolution could play a role in improving redshift determinability.

@@

The theoretical redshift distribution of sGRBs depends on the type of delay-time
function used to model the progenitor system. The likelihood preferred lognormal
time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
$z = 0.9$. According to \citet{Wanderman2015}, this preference depends
critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The higher
determined redshift of GRB~111117A, and the likely number of additional high-z
sGRB could change the preferred time delay models. The redshift of GRB~111117A,
on the other hand, is close to the expected peak in sGRB rate calculated using
the power law delay time models \citep{Behroozi2014, Wanderman2015,
	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.

@@

Using the inferred stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. To investigate
the possible impact of the presence of an old stellar population, we followed
\citet{Papovich2001} and re-fitted the observed SED with the best-fit template
to which an additional stellar population of old stars was added. For each
galaxy, this old population was set as the SPSs with same parameters that the
best-fit SED excepts the age, which was set to the age of the Universe at the
observed redshift. In principle, this can constrain the maximum contribution of
old populations within the photometric error bars \citep[see][ for
details]{Papovich2001}. We find a negligible contribution to the stellar mass
(i.e. variations much smaller than the statistical uncertainty associated with
the best-fit template). The delay time between formation and explosion is well
accommodated by the models of \citet{Belczynski2006}, although the longest delay
times are excluded. This is especially true given the late type nature of the
host \citep{OShaughnessy2008}.

@@

Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
upper end of the sGRB host SFR distribution and no significant amount of dust is
present. The high $N_H$ is difficult to reconcile with the large projected host
offset and the absence of dust. One possible explanation could be, that
GRB~111117A is formed through a prompt channel of sGRB formation and originates
in a star forming region located in the outskirts of the host.

@@

\begin{acknowledgements}
We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event. We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
%	
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
%
SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
%
JJ acknowledges support from NOVA and NWO-FAPESP grant for advanced
instrumentation in astronomy.
%
NRT and KW acknowledge support from STFC Consolidated
Grant ST/N000757/1
%
CT acknowledges support from a Spanish National Research Grant of Excellence
under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
fellowship under grant number RyC-2012-09984.
%
AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy
and Competitiveness through project AYA2014-58381-P. Partly based on
observations made with the Gran Telescopio Canarias (GTC).
%
ZC acknowledges support from the Spanish research project AYA 2014-58381-P and
support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
%
RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
%
This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}. The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
%
This work made use of observations obtained with the Italian 3.6 m Telescopio Nazionale Galileo (TNG) operated on the island of La Palma by the Fundaci\'on Galileo Galilei of the INAF (Istituto Nazionale di Astrofisica) at the Spanish Observatorio del Roque de los Muchachos of the Instituto de Astrof\'isica de Canarias.
%
Based on data from the GTC Archive at CAB (INTA-CSIC) and on observations obtained at the Gemini Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnología e Innovación Productiva (Argentina), and Ministério da Ciência, Tecnologia e Inovação (Brazil).

@@

\bibliographystyle{aa_arxiv}
\bibliography{GRB111117A}

>>>>

\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}

@@

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citet{Kouveliotou1993},
where the duration T$_{90} < 2$ has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

@@

The electromagnetic signals from sGRBs are a promising channel to accurately
localize NS mergers, which holds the promise for a detection of an associated
gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
of a sGRB and a GW will provide new a promising ways to constrain the binary
inclination angle \citep{Arun2014} and measure cosmological distances
\citep{Nissanke2010}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum. The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010} run in physical
mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise ratio of the continuum in the near-infrared arm is too low to
use the optimal extraction algorithm \citep{Horne1986} and therefore the
extraction is carried out with a simple aperture covering the entire host, where
we sum the flux in the aperture and derive the error by summing the
corresponding pixel errors in quadrature. An overview of the spectroscopic
observations is given in Table~\ref{tab:spec_overview}, and the slit position is
shown in Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in
Fig.~\ref{fig:SED}.

@@

The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
profiles shows an asymmetric "double-horned" profile, indicating that we are
seeing a galaxy with a large degree of coherent rotational motion relative to
the line-of-sight. If we assume that we are viewing a spiral galaxy edge-on,
this is a measure of the rotational velocity of the gas. If we assume that the
spectral resolution and the turbulent width of the lines are negligible compared
to the rotational velocity, we can, based on the projected size of the source
and the width of the lines, put a constraint on the dynamical mass of the galaxy
\citep{DeBlok2014}. Based on the physical size along the slit and the velocity
width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
M$_\odot$. Because we are viewing the host inclined at an angle relative to
edge-on and because the slit is not aligned along the long axis of the host,
this value is an upper limit.

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, there
is likely no afterglow contamination on the measured photometry.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X-ray temporal and spectral analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The redshift evolution of $N_H$ in the hosts of lGRBs is not reproduced by
\citet{Buchner2017}, using a different $N_H$ inference methodology. Instead a
correlation between $N_H$ and host stellar mass is suggested. Assuming that the
different $N_H$-fitting methodologies yield comparable results, GRB~111117A is
an outlier from the relation suggested by \citet{Buchner2017} by more than the
intrinsic scatter, although some lGRB hosts populate a similar region in the
$N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
surface luminosity at the explosion site \citep{Lyman2017}. The large offset of
GRB~111117A relative to the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013} is difficult to reconcile with galaxy-scale gas providing
the X-ray absorption. Along with the absence of dust, the large offset from the
host center indicates that the high $N_H$ arises because the density in the GRB
surrounding is high, or because the light from the afterglow traverses a region
of dense gas.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs. The detections are replaced with contours and the limits are shown with arrows.}
	\label{fig:NH_z}
\end{figure}

@@

If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed 60
per cent of the redshifts (6 out of 10 hosts), were they at the redshift of
GRB~111117A due to the hosts being too faint. The corresponding number is around
45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
the average SFR of galaxies hosting lGRBs is higher than for galaxies hosting
sGRBs, the fraction of missed burst redshifts is likely higher although the
cosmic SFR evolution could play a role in improving redshift determinability.

@@

The theoretical redshift distribution of sGRBs depends on the type of delay-time
function used to model the progenitor system. The likelihood preferred lognormal
time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
$z = 0.9$. According to \citet{Wanderman2015}, this preference depends
critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The higher
determined redshift of GRB~111117A, and the likely number of additional high-z
sGRB could change the preferred time delay models. The redshift of GRB~111117A,
on the other hand, is close to the expected peak in sGRB rate calculated using
the power law delay time models \citep{Behroozi2014, Wanderman2015}.

@@

Using the inferred stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. The delay time
between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest delay times are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.

@@

Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
upper end of the sGRB host SFR distribution and no dust is present. The high
$N_H$ is difficult to reconcile with the large projected host offset and the
absence of dust. One possible explanation could be, that GRB~111117A is formed
through a prompt channel of sGRB formation and originates in a star forming
region located in the outskirts of the host.

@@

\begin{acknowledgements}
We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event. 
%	
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
% 
JJ acknowledges support from NOVA and NWO-FAPESP grant for advanced
instrumentation in astronomy.
%
NRT and KW acknowledge support from STFC Consolidated
Grant ST/N000757/1
%
CT acknowledges support from a Spanish National Research Grant of Excellence
under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
fellowship under grant number RyC-2012-09984.
%
AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy
and Competitiveness through project AYA2014-58381-P. Partly based on
observations made with the Gran Telescopio Canarias (GTC).
%
ZC acknowledges support from the Spanish research project AYA 2014-58381-P and
support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
%
RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
%
This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}. The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
%
This work made use of observations obtained with the Italian 3.6 m Telescopio Nazionale Galileo (TNG) operated on the island of La Palma by the Fundaci\'on Galileo Galilei of the INAF (Istituto Nazionale di Astrofisica) at the Spanish Observatorio del Roque de los Muchachos of the Instituto de Astrof\'isica de Canarias.
%
Based on data from the GTC Archive at CAB (INTA-CSIC) and on observations obtained at the Gemini Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnología e Innovación Productiva (Argentina), and Ministério da Ciência, Tecnologia e Inovação (Brazil).

@@

\bibliographystyle{mnras}
\bibliography{GRB111117A}
